---
title: "Exercise 2 - Data Visualisation"
author: "Keaan Amin"
date: '`r format(Sys.time(), "%d %b %Y")`'
output: pdf_document
---

Grab the data.csv file from the Github Directory (https://github.com/keaan95/virtual-elective/tree/master/Week1) and put it into your Working Directory.


# EXERCISE 1
## 1.1. Add the data file to your ENVIRONMENT.
TIP - Remember to Set your Working Directory to the location of your data file.
  
```{r}
# You can fetch your WORKING directory as follows
getwd()
# Change to your local directory - in my case it is as follows:
setwd("C:/Users/keaan/OneDrive - Newcastle University/HLA - Virtual Elective/Week1/")
df <- read.csv("data.csv")
```

# EXERCISE 2
## 2.1. Visualise the first five columns and rows.

```{r}
df[1:5,1:5]
```

## 2.2. How many patients are in the data?

```{r,eval=FALSE}
nrow (df)
```

## 2.3. Let's Plot Radius against Perimeter Means of Our Patients

```{r}
plot(df$radius_mean,df$texture_mean,
     xlab="Ex")
```

## 2.3. Let's Add some Legends and a Title

```{r}
plot(df$radius_mean,df$texture_mean,
     xlab="Radius Means",
     ylab="Texture Means",
     main="Breast Cancer Patients of Radius against Texture Means")
```


## 2.4. Let's Colour By Malignant vs. Benign Patients

In order to do this we need to change our column 'data type'

```{r,eval=FALSE}
# View the data type of structures of our dataframe 'df'
str(df)
```

Our 'df$diagnosis' are a bunch of  'M's or 'B's - known as 'characters.' We need this data type to be a 'factor' so R recognises it belongs to a particular group. We can do this by:

```{r,eval=FALSE}
as.factor(df$diagnosis)
```

Using the col variable, we can specify this:

```{r}
plot(df$radius_mean,df$texture_mean,
     col=as.factor(df$diagnosis),
     xlab="Radius Means",
     ylab="Texture Means",
     main="Breast Cancer Patients of Radius against Texture Means")
legend("topright",legend = c("M", "B"), pch=c(1,1), col=c("red","black"))
```

